\section{Introduction}
\label{sec:introduction}
This report describes the efforts in defining a concrete method of extending the Futhark Programming Language with a module system.

A module system is here defined as a programming feature, which allows the programmer to package functionality into a discrete code package.

A module can then be shared and utilized in any other program, by including it in the source code. Widely known modules include \texttt{NumPy}\footnote{\url{http://www.numpy.org/}}, the Python package for
data modelling and computing, \texttt{NLTK}\footnote{\url{http://nltk.org}},
the Natural Language Toolkit (also for Python), or \texttt{leftpad}\footnote{\url{https://www.npmjs.com/package/left-pad}} for node.js.


A well written module can become very popular. In some cases, a programmers choice of programming language for a given project, can be decided by the number and quality of available, project relevant modules.


The module system design for Futhark has been inspired by the module system implemented in Standard ML\cite{crash_course}. 

Not only does the Standard ML modules enable structures with predefined implementations. It also defines a system of \textit{functors\footnote{Not to be confused with neither mathematical functors or functors from Haskell}}, which defines abstract implementations of structures.


These abstract structures are applied on an argument of a fitting signature, returning a realized structure. This makes it possible to write structures that does not commit to one certain type, but instead inherits the functionality defined in the given argument.


This spares the developers for writing a lot of almost identical code. Refer to (Figure 2\ref{first_functor_example}) for a simple example.


\textbf{Reader expectations}:
The reader is expected to have experience with statically typed languages, functional programming, and preferable the Haskell programming language.
\pagebreak
\subsection{Motivation}
\label{subsec:label}
The Futhark language a young language\footnote{It's official web site \url{http://Futhark-lang.org} was launched in March 2016.}, and is in continous developement. Extending this language to support modules would increase the usability of Futhark, which could be a factor in expanding the user base and user retention
of Futhark.


The following subsections contain the features this project is implementing, and examples of why these features are desirable.
\subsubsection{Abstraction increases readability.} 
\begin{figure}
\begin{mdframed}
\begin{lstlisting}
fun {f32, {f32, f32, f32}, {f32, f32, f32}} multiply_velocity(
	{f32 r, 
	{f32 x_pos , f32 y_pos , f32 z_pos} , 
	{f32 x_dir,f32 y_dir, f32 z_dir}} , f32 k) 
	  =
    {r, {x_pos, y_pos, z_pos}, {k * x_dir , k * y_dir, k * z_dir }}
\end{lstlisting}
\caption{A function polluted by all the references to \texttt{f32} \label{noalias}}
\end{mdframed}
\label{noaliasing}
\end{figure}
Type aliasing lets us abstract from the actual definition of data types.
When writing our source code, we can define our type aliases before writing the
rest of our program. Take for example a sphere in three dimensional space:
It has \textbf{1) a radius}, \textbf{2) a position} , and \textbf{3) a direction} that it is
moving in.\\
Let us define a function that multiplies the speed of the sphere by a factor k: (Figure ~\ref{noalias})
\clearpage

With type aliasing, we can compartmentalize the data type, and remove the need
for explicitally typing out every parameter of function input.
Coupled with helper functions, we can now multiply the speed of the sphere (Figure ~\ref{sphere_example}) without being explicit about the actual types at each function call.
\begin{figure}
\begin{verbatim}
type vec3 = {f32, f32, f32}
type position, direction = vec3 
type radius = f32
type sphere = {radius, position, direction}

fun vec3 multiply_vector(vec3 {pos_x, pos_y, pos_z}, f32 k) =
  {k * pos_x , k * pos_y , k * pos_z}

fun sphere multiply_velocity(
  sphere {radius, position, direction}, f32 factor)  =
    let new_direction = multiply_vector(direction, factor)
    in {radius, position, new_direction}
\end{verbatim}
\caption{Type aliasing abstracts type definitions into names that are more
  fitting for the function declaration at hand \label{sphere_example}}
\end{figure}

\noindent As we are now using the sphere type as the function argument, we can pattern
match on the type aliased values contained in the sphere type.
Most importantly, the vectors of the sphere are abstracized into a single variables
instead of tuples.
\\
The end result is a shorter, more readable program.

\subsubsection{Compartmentalization of functionality increases usability:}
Splitting code functionality into multiple files, will allow the programmer to
compile and type check these modules individually. The programmer can edit and
contribute to these files independently of the programs, which includes these
modules.

\subsubsection{Approximating higher order functionality whilst keeping performance}
\label{subsec:higherorderfunctionality}
It is possible to express higher-order functionality in Futhark, without taking
a performance hit in the compiled Futhark code.
%% reference to slow higher order functionality
We will reiterate on the concept of structures (Section \ref{sec:structures}), by introducing
the concept of functors.\\
First we repeat the three-dimensional vector structure
from earlier, but without declaring any particular
primitive type (Section ~\ref{Futharks_types}) as the contained type of the vector (Figure ~\ref{struct_vec3}).
\begin{figure}
  \begin{verbatim}
  functor Vec3 ( number_type : NumType) { 
        type t = number_type.t
        type vector = {t, t, t}
        fun vector add( ... ) = ...
        fun vector subtract( ... ) = ...
        fun vector multiply ( ... ) = ...
        fun vector divide ( ... ) = ...
    }
  \end{verbatim}
  \caption{A functor defining a vector structure \label{struct_vec3}}
\end{figure}

The structure (Figure ~\ref{struct_vec3}) above cannot be used on its own. Type \texttt{t} is not
instantiated, and the module cannot be type checked, which causes an error.


We can solve that problem, by applying the abstract structure on a concrete
structure.

\begin{figure}
\begin{verbatim}
  struct Int {
    type t = int 
  } : NumType
  struct IntVec3 = Vec3(Int)
\end{verbatim}
\caption{The IntVec3 structure is an instance of Vec3. \label{first_functor_example}}
\end{figure}
We can now access the structure \texttt{IntVec3} throughout the rest of the
program.

The structure \texttt{IntVec3} is \texttt{Vec3}, except all instances of type \texttt{t} in \texttt{Vec3} is exchanged with type \texttt{int}.


\textbf{To recap:} functors allows us to define an abstract implementation of some structure \textbf{ONCE}, and lets us instantiate this structure any number of times, each time with our own type.


From a performance-concerned point of view, the module system is desirable, as the Futhark compiler handles all used and included structures at compile time.
\subsection{Problem definition}
\label{subsec:probdef}
Is it possible to implement a module system in Futhark, which displays features comparable to the module system implemented in Standard ML?

\subsubsection{Scope of project}
\label{subsec:project_scope}
The scope of the project is
\begin{itemize}
  \item to define and implement a type aliasing system in Futhark
  \item to define and implement a module system in Futhark, which has:
    \begin{itemize}
      \item The definition of structures containing types and functions
      \item Nested modules; meaning that any structure can contain a structure
      \item A well defined way of referring to types, structures and functions
        contained in a structure
    \end{itemize}
  \item to research the possibility of implementing functor and signature functionality, so that Futhark supports the definition of abstract structures and concretizations of these structures. A suggested design for the implementation of functors should be part of this project.
\end{itemize}
\clearpage

\subsection{Related work}
Futhark modules are designed with Standard MLs module system in mind. Thus, 
\textit{The Definition of Standard ML '97}\cite{sml_book} has been a source of inspiration
for the behaviour of the Futhark module system. \\
However, I have not read the code implementation of SML modules. My code supervisor and I did not find it necessary to lean on the SML implementation of modules:\\
The Futhark compiler could readily support extensions to the language, and it was not difficult to
figure out a way to implement the module system.
\\
\\
Implementing SML modules in other languages is not a unique idea:\\
Yaraw Amin\cite{scala_modules} has shown a way of implementing SML modules in Scala
language. However, he was able to exploit functionality that was already available in
Scala, to create his module system.
\\
\\
Mainly, he was able to use Scala \texttt{functions} as functors to create new Scala
\texttt{classes} (comparable to structures), just as Scala
\texttt{traits} were used for implementing signatures.
\\
\\
Futhark had neither functions (for structures), structures or signatures at the
beginning of the project, which made these necessary to add to the compiler itself. Therefore, the following project is not necessarily original research (as modules are a well known concept,) but an original implementation in Futhark. Original as opposed to i.e. a port of the implementation from SML/NJ.